SSA based global code motion of identical computations

### 2017-04-14 hoist.tex
remove partial hoist

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline when their code size decreases because inliner heuristics in \LLVM{}
  depends on instruction count.
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting/sinking expressions out of branches, code hoisting can effectively
  reduce the amount of code to be speculatively executed and can reduce the
  critical path;
\item to reduce interference/register pressure: For example if all the use
  operands of an instruction (I) of the form a = b op c are last usages (a kill)
  then hoisting will reduce the interference by two but increase by one across
  the path I gets hoisted.
\item for SIMD architectures, which execute all branches, it will reduce the
  total number of instructions to be executed.
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted or sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

@@

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC and AVAIL value expressions computed
for \PRE{}.  The algorithm hoists top down in a predecessor where the value is in
ANTIC.  It uses ANTIC[B] to know what expressions will be computed on every
path from B to exit, and can be computed in B: that is the safety condition.  It
uses AVAIL[B] to subtract out those values already being computed (because they
are already available): this is a redundancy condition.  The cost function is:
for each hoist candidate, if all successors of B are dominated by B, then we
know insertion into B will eliminate all the remaining computations.  It then
checks to see if at least one successor of B has the value available.  This
avoids hoisting it way up the chain to ANTIC.  It also checks to ensure that B
has multiple successors, since hoisting in a straight line is pointless.  The
algorithm continues on down the dominator tree, iterating with \PRE{} until no more
changes.  One advantage of \GCC{} implementation is that it works in sync with the
\GVN{}-\PRE{} such that when new hoisting opportunities are created by \GVN{}-\PRE{},
code-hoisting will hoist them.

@@

When aggressively for code-size (at optimization level -Oz), we allow hoisting
of scalars even when the expression is not very busy. This creates unnecessary
computation on paths where the expression is not needed so the program may run slow
but reduces the code size. TODO: Experimental results.

@@

>>>>

\begin{itemize}
\item to reduce the code size of the program;
\item to improve function inlining heuristics: functions become cheaper to
  inline when their code size decreases because inliner heuristics in \LLVM{}
  depends on instruction count.
\item to expose more instruction level parallelism: by hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles;
\item to help out-of-order processors with speculative execution of branches: by
  hoisting/sinking expressions out of branches, code hoisting can effectively
  reduce the amount of code to be speculatively executed and hence reduce the
  critical path;
\item to reduce interference/register pressure: For example if all the use
  operands of an instruction (I) of the form a = b op c are last usages (a kill)
  then hoisting will reduce the interference by two but increase by one across
  the path I gets hoisted.
\item for SIMD architectures, which execute all branches, it will reduce the
  total number of instructions to be executed.
\item to improve passes that do not work well with branches:
  \begin{itemize}
  \item to improve loop vectorization by reducing a loop with control flow to a
    loop with a single BB, should all the instructions in a conditional get
    hoisted or sinked;
  \item to enable more loop invariant code motion (LICM): as LICM does not
    reason about instructions in the context of loops with conditional branches,
    code-hoisting is needed to move instructions out of conditional expressions
    and expose them to LICM.
  \end{itemize}
\end{itemize}

@@

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions at
basic block B i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
top down to a predecessor.  It uses to know what expressions will be computed on
every path from B to exit, and can be computed in B: that is the safety
condition.  It uses AVAIL\_OUT[B] to subtract out those values already being
computed.  The cost function is: for each hoist candidate, if all successors of
B are dominated by B, then we know insertion into B will eliminate all the
remaining computations.  It then checks to see if at least one successor of B
has the value available.  This avoids hoisting it way up the chain to ANTIC.  It
also checks to ensure that B has multiple successors, since hoisting in a
straight line is pointless.  The algorithm continues top down the dominator
tree, iterating with \PRE{} until no more changes.  One advantage of \GCC{}
implementation is that it works in sync with the \GVN{}-\PRE{} such that when
new hoisting opportunities are created by \GVN{}-\PRE{}, code-hoisting will
hoist them.

@@

@@

Since stores and calls are hoisted the least \ref{table:experimental-results}
the performance does not change much whether they are hoisted or not.
